### 2020-11-24 presentation/spec.tex

\section{Cipher Specifications}

\begin{frame}{Slide One}

\end{frame}

\begin{frame}{Slide Two}

\end{frame}

\begin{frame}{Slide Three}

\end{frame}