### 2020-11-27 presentation/conclu.tex

\section{Conclusion}

\begin{frame}{Conclusion}
  \begin{beamerboxesrounded}{Similarity with AES}
    SQUARE, which is a predecessor of AES is very similar to AES in it's structure and S-box. And shares some common attacks.
  \end{beamerboxesrounded}
  \vspace{5mm}
  \begin{beamerboxesrounded}{Attacks}
    Practical attacks for up to six rounds are known for SQUARE  and hence the number rounds is 8 following a conservative approach.
  \end{beamerboxesrounded}
  \vspace{5mm}
  \begin{beamerboxesrounded}{Use in real world}
    Even though the known full round attacks are not practical, the authors recommend against using it in applications due to lack of intense public scrutiny.
  \end{beamerboxesrounded}
\end{frame}