### 2020-11-28 presentation/intro.tex
fix latex errors

\section{Introduction}

\begin{frame}{SQUARE Cipher}
  \begin{beamerboxesrounded}{Introduction}
    Square is an iterated block cipher. Block length and key length is 128 bits.
    The original design of Square concentrates on the resistance against differential and linear cryptanalysis. However, after integral attack cipher rounds were extended to eight rounds. Now,There are total eight rounds in SQUARE cipher.And the round transformation of Square is composed of four distinct transformations($\theta$,$\gamma$,$\pi$,$\sigma$).
  \end{beamerboxesrounded}
\end{frame}